---
title: "MinimalDefaultAmount"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Set Working Directory}
setwd("F:/MSCSNyuPoly/Fall2016/FoundationsOfDataScience/Project/Work")
```

install.packages("sqldf")

data2015 <- read.csv("2015_dataMain.csv")
#COnvert factor to character
data2015$Total.Violation.Amount <- as.character(levels(data2015$Total.Violation.Amount)[data2015$Total.Violation.Amount])
#Replace $ sign with blank
data2015$Total.Violation.Amount <- sapply(data2015["Total.Violation.Amount"], function(x) as.character(gsub("\\$", "", x)))
#Convert character to numeric
data2015$Total.Violation.Amount <- as.numeric(data2015$Total.Violation.Amount)

#Select only Total Violation Amount column
data2015Amt <- subset(data2015, select = c("Total.Violation.Amount"))
#Total number of violations
ttlViolation <- nrow(data2015Amt)
#Aggregation to find percentage occurenece of each penalty amount
options(scipen = 999)  #To disable scietific notation
ttlAgg <- do.call(data.frame, aggregate(cbind(count = Total.Violation.Amount) ~ Total.Violation.Amount, data = data2015Amt, 
          FUN = function(x) c({round((NROW(x) / ttlViolation) * 100, digits = 4)}, {NROW(x)})))
colnames(ttlAgg) <- c("Total.Violation.Amount", "ttlPercent", "ttlCnt")

#Select data where Hearing Status is PAID IN FULL and Decision Date is blank. It means that amount has been paid in full before taking any decision
data2015Paid <- subset(data2015, Hearing.Status == "PAID IN FULL" & Decision.Date == "", select = c("Total.Violation.Amount"))
#Aggregation  to find percentage occurence of penalty where user has paid fully
paidAgg <- do.call(data.frame, aggregate(cbind(count = Total.Violation.Amount) ~ Total.Violation.Amount, data = data2015Paid, FUN = function(x)c({round((NROW(x) / ttlViolation) * 100, digits = 4)}, {NROW(x)})))
colnames(paidAgg) <- c("Total.Violation.Amount", "paidPercentFromTotal", "paidCount")

#Left join of ttl Percentage and paid percentage
mergeData <- merge(x = ttlAgg, y = paidAgg, by = "Total.Violation.Amount", all.x = TRUE)
#Replace NA with 0
mergeData <- data.frame(apply(mergeData, 2, function(x){replace(x, is.na(x), 0)}))
#Calculate percentage violations per amount where penalty was paid before decision
mergeData$paidPercentPerPenalty <- (mergeData[,"paidCount"] / mergeData[,"ttlCnt"]) * 100

mergeData
colnames(mergeData)
#Plot penalty amount vs percentage of violations in which penalty was paid before decision

sampleMergeData <- mergeData[mergeData$Total.Violation.Amount < 4000,]


plot(mergeData[,"Total.Violation.Amount"], mergeData[,"paidPercentPerPenalty"], type = 'o', main = "No of times Penalty Paid before Decision per Amount", xlab = "Penalty Amount", ylab = "Percentage", col = "blue")

#Same plot with less data
plot(sampleMergeData[,"Total.Violation.Amount"], sampleMergeData[,"paidPercentPerPenalty"], type = 'o', main = "No of times Penalty Paid before Decision per Amount", xlab = "Penalty Amount", ylab = "Percentage", col = "blue")

sampleMergeData <- mergeData[mergeData$Total.Violation.Amount < 1000,]
#Plot penalty amount vs ttl percentage and penalty paid percentage
plot(sampleMergeData[,"Total.Violation.Amount"], sampleMergeData[,"ttlPercent"], type = 'o', main = "", xlab = "Penalty Amount", ylab = "Percentage", col = "blue")
lines(sampleMergeData[,"Total.Violation.Amount"], sampleMergeData[,"paidPercentFromTotal"], col = "red")

#Barplot
head(mergeData)
rownames(sampleMergeData) <- sampleMergeData$Total.Violation.Amount
barplot(t(as.matrix(sampleMergeData[,c(2,4)])), col=c("darkblue","red"), legend=c("Total", "Paid"),ylim=c(0, 60), xlab = "Amount", main = "Per Amount Paid Penalty Before Decision", beside=TRUE)

#We can see that for lower penalty amount there is more percentage to pay penalty beofre decison. It is highest for penalty amount $25.


#How much revenue is lost on defaulted penalties
copyData2015 <- data2015
Convert violation date to month
#Extract month from violation date
copyData2015$Violation.Date <- format(as.Date(as.character(copyData2015$Violation.Date), "%m/%d/%Y"), "%m")
#Select defaulted violations
data15MonthDefaulted <- subset(copyData2015, (Hearing.Status == "DOCKETED" | Hearing.Status == "DEFAULTED") & Hearing.Result == "DEFAULTED", select = c("Violation.Date", "Total.Violation.Amount"))

#Select subset of data for all violations
data15Month <- subset(copyData2015, select = c("Violation.Date", "Total.Violation.Amount"))

#Aggregate default violations
dataMonthDeafaultedAgg <- do.call(data.frame, aggregate(cbind(sum = Total.Violation.Amount) ~ Violation.Date, data = data15MonthDefaulted, FUN = sum))
colnames(dataMonthDeafaultedAgg) <- c("Violation.Date", "DefaultedSum")
#Aggregate all violations
data15MonthAgg <- do.call(data.frame, aggregate(cbind(sum = Total.Violation.Amount) ~ Violation.Date, data = data15Month, FUN = sum))
colnames(data15MonthAgg) <- c("Violation.Date", "ttlSum")
#Join and calculate percentage
mergeData <- merge(x = data15MonthAgg, y = dataMonthDeafaultedAgg, by = "Violation.Date", all.x = TRUE)
mergeData$Violation.Date <- as.character(mergeData$Violation.Date)

#Plot bargraph
rownames(mergeData) <- mergeData$Violation.Date
barplot(t(as.matrix(mergeData[,c(2,3)])), col=c("darkblue","red"), legend=c("Total", "Defaulted"), xlab = "Month", main = "Monthwise Total Revenue and lost revenue", beside=TRUE)



#Find percentage of defaulted violations per agency, violationLoc, Penalty, Respondent Location
copyData2015 <- data2015
colnames(data2015)
#Select defaulted violations
data15Defaulted <- subset(copyData2015, (Hearing.Status == "DOCKETED" | Hearing.Status == "DEFAULTED") & Hearing.Result == "DEFAULTED", select = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"))

#Select subset of data for all violations
data15All <- subset(copyData2015, select = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"))

#Aggregate default violations
dataDeafaultedAgg <- do.call(data.frame, aggregate(cbind(DefaultedCount = Total.Violation.Amount) ~ (Charge..1..Code.Description + Violation.Location..Zip.Code. + Respondent.Address..Zip.Code. + Total.Violation.Amount), data = data15Defaulted, FUN = function(x)c({NROW(x)})))

#Aggregate all violations
data15AllAgg <- do.call(data.frame, aggregate(cbind(TotalCount = Total.Violation.Amount) ~ (Charge..1..Code.Description + Violation.Location..Zip.Code. + Respondent.Address..Zip.Code. + Total.Violation.Amount), data = data15All, FUN = function(x)c({NROW(x)})))
colnames(data15AllAgg)
#Join and calculate percentage
mergeData <- merge(x = data15AllAgg, y = dataDeafaultedAgg, by = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"), all.x = TRUE)
mergeData$Percentage <- (mergeData[,"DefaultedCount"]/mergeData[,"TotalCount"])*100
mergeData <- data.frame(apply(mergeData, 2, function(x){replace(x, is.na(x), 0)}))
mergeData$Percentage <- as.numeric(as.character(mergeData$Percentage))

mergeData <- subset(mergeData, as.numeric(TotalCount) > 100,select = c( "Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount", "TotalCount", "DefaultedCount", "Percentage"))
#Top 100 combination of Issuing Agency, Violation Address, Respondent Address, Total Violation amount where most of violations are defaulted
top100 <- head(mergeData[order(mergeData$Percentage,decreasing=T),], 100)
bottom100 <- head(mergeData[order(mergeData$Percentage),], 100)




